The Multi-Stencil Framework: A Component-Based Approach to Orchestrate Stencils

### 2017-03-06 ARTICLES/IJPP/articleR1/parallelism.tex
corrections parallelism.tex + evaluation.tex

\medskip
More formally, the data parallelization of a multi-stencil program 
\begin{equation*}
\mathcal{MSP}(\mathcal{M},\Phi,\mathcal{D},\mathcal{N},\Delta, \mathcal{S},T,\Gamma)
\end{equation*}
consists in a partitioning of the mesh $\mathcal{M}$ in $p$ sub-meshes $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
This step can be performed by an external graph partitioner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is addressed by the third party DDS component instantiated from GA.

@@

However, as a multi-stencil program is an iterative program, computations which happen after $k_j$ at the time iteration $t$ have also been computed before $k_j$ at the previous time iteration $t-1$. For this reason another case of synchronization has to be defined.

@@

One can notice that both types of synchronizations are performed by all resources.

\begin{mydef}
The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[k_i]_{0 \leq i \leq n-1}$ and $l_2=[k'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[k_0,\dots,k_{n-1},k'_0,\dots,k'_{m-1}]$.
\end{mydef}

@@

\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels can be extracted:
\begin{equation*}
\Gamma = [k_0,k_1,k_2,k_3,k_4,k_5,k_6,k_7,k_8]
\end{equation*}
From this ordered list of computation kernels $\Gamma$, and from the rest of the multi-stencil program, synchronizations can be automatically detected from the call to $F_{sync}(\Gamma,0)$ to get the synchronized ordered list of kernels:
\begin{equation}
\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_5,k_6,k_7,k_{7;8}^{sync},k_8],
\label{eq:exsync}
\end{equation}
\begin{subequations}
where
\begin{align}
        k_{0;1}^{sync}=(\emptyset,\{(B,nce)\},(B,\cup_{\phi \in D_1} nce(\phi))),\\
        k_{1;4}^{sync}=(\emptyset,\{(C,nec)\},(C,\cup_{\phi \in D_4} nec(\phi))),\\
        k_{7;8}^{sync}=(\emptyset,\{(I,ncc)\},(I,\cup_{\phi \in D_8} ncc(\phi))).
\end{align}
\end{subequations}

@@

Those definitions are known as \emph{data hazards classification}. However, a specific condition on the computation domain, due to the specific domain of multi-stencils, is introduced for the write after write case. One can note that the \texttt{independent} keyword of Fig.~\ref{fig:grammar} is useful in this case.

@@

One can notice that $\Gamma_{dep}$ is the dependency graph of the computations of a multi-stencil program, but it only takes into account a single time iteration. A complete dependency graph of the simulation could be built. This is a possible extension of this work.

@@

Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or sub-graph, called $N$, such that the following property is verified :

\begin{myth}
The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as an induced sub-graph.
\end{myth}

As $\Gamma_{dep}$ is a DAG, by removing N-Shapes it is transformed to a MSPD. The intuition is illustrated in Fig.~\ref{fig:over}. Considering the figure without the dashed line, the sub graph forms a "N" shape. The fact is that this shape cannot be represented as a composition of sequences or parallel executions.
To remove such forbidden N-shapes of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag as illustrated in Figure~\ref{fig:over}. By adding this arc to the DAG, it is possible to identify its execution as $sequence(parallel(k_0;k_2);parallel(k_1;k3))$ represented by the TSP tree of Fig.~\ref{fig:solu}.

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (bc0) at (-1,0) {};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (0,-1) {$k_2$};
   \node[] (bc2) at (-1,-1) {};
   \node[] (c3) at (2,-1) {$k_3$};

  \path[->]
    (bc0) edge [dotted] node {} (c0)
    (bc2) edge [dotted] node {} (c2)
    (c0) edge node {} (c1)
          edge [dashed] node [swap] {} (c3)
    (c2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
\caption{Over-constraint on the forbidden $N$ shape.}
\label{fig:over}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (S0) at (0,0) {$\mathcal{S}$};
   \node[] (bS0) at (-1,-1) {};
   \node[] (P1) at (-1,1) {$\mathcal{P}$};
   \node[] (aP1) at (-1.5,2) {$k_0$};
   \node[] (aaP1) at (-0.5,2) {$k_2$};
   \node[] (P2) at (1,1) {$\mathcal{P}$};
   \node[] (c1) at (0.5,2) {$k_1$};
   \node[] (c3) at (1.5,2) {$k_3$};

  \path[->]
    (bS0) edge [dotted] node {} (S0)
    (S0) edge node {} (P1)
         edge node {} (P2)
    (P1) edge node [swap] {} (aP1)
          edge node [swap] {} (aaP1)
    (P2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
\caption{TSP tree of Fig.~\ref{fig:over}.}
\label{fig:solu}
\end{center}
\end{figure}

@@

Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization. Even though the MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations, building the $TSP$ tree makes available an easy optimization when the data parallelization technique is the only one used. 
This optimization consists in proposing a valid merge of some computation kernels inside a single space loop. This is called a fusion. As previously explained in Section~\ref{sect:formalism}, MSL restrict the definition of a numerical computation by writing a single quantity at a time which is more natural for numericians, and which avoid errors in manual fusion or counter-productive fusions for task parallelization. MSF guarantee that proposed fusions are correct and will not cause errors in the final results of the simulation.

@@

The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where the $parent(k)$ function returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$. Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$ otherwise.

@@

Thirs, and similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one. The construction of the tree also guarantees that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$). Thus, $w_i$ the quantity written by $k_i$ is common to these computations. Considering the following:
\begin{itemize}
\item $D_i \neq D_j$, which means that loop fusion is by default not possible,
\item $(r_j,n_j)$ is the pair read by $k_j$ for which $r_j=w_i$ and for which $n_j:D_j \rightarrow D_i^m$
\end{itemize}
the fusion of $k_i$ and $k_j$ is possible if and only if $\exists n:D_i \rightarrow D_j \in \mathcal(N)$ such that
\begin{equation*}
\bigcup_{\phi \in D_i} n(\phi) = D_j
\end{equation*}
This means that even if domains are different, a loop fusion is possible if an adequate neighborhood function can be found. One can note that this particular fusion case is equivalent to a \emph{scatter} optimization, often used when using unstructured meshes. One can also note that the computation $k_j$ will be written in a different manner if a scatter fusion is performed or not. This is not a problem by using MSF as the fusion is proposed before the developer actually write the numerical code of $k_j$. The developer will be notified of fusions in the output of MSF. This particular case is illustrated in Fig.~\ref{fig:fus3}.

@@

@@

MSC takes a MSL file written following the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly of components represented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation. In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill-in the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
The overall behavior of the compiler is as follows: %presented in Algorithm~\ref{alg:compiler}.
\begin{enumerate}
 \item it parses the MSL input file and generates $\Gamma$, the list of computation kernels,
 \item from $\Gamma$, it builds $\Gamma_{sync}$, the list including synchronizations for data parallelism using Algorithm $F_{sync}$ introduced in Section~\ref{sect:parallelism},
 \item from $\Gamma_{sync}$, it builds $\Gamma_{dep}$, the DAG supporting hybrid parallelism using Algorithm $F_{dep}$ introduced in Section~\ref{sect:parallelism},
 \item it then removes the N-Shapes from $\Gamma_{dep}$ to get a MSPD graph, and generates its serie-parallel binary tree decomposition $TSP$,
 \item it performs the fusion of kernels in $TSP$ if required (data parallelization only),
 \item it transforms GA to generate its output specialized component assembly.
\end{enumerate}

The last step of this compilation process is detailed below. It is composed of four steps:
\begin{enumerate}
 \item it instantiates $DDS$ and $Data$ components by using components implemented by a third party HPC specialist,
 \item it generates $K$ components responsible for each computation kernel of the simulation,
 \item it generates a new $Scheduler$ component,
 \item it replaces $Computations$ component by a generated sub-assembly that matches $TSP$ by using $Scheduler$, $K$ and $Sync$ components.
 \end{enumerate}

% detailed each component (interfaces)
New components have been introduced above and need to be explained. A $K$ component is a component into which the developer will write numerical code. It could represents a single computation kernel described by the numerician using MSL, or it could represents the fusion of multiple computation kernels. In any case the name of the generated component will use kernel identifiers used in the MSL description. A $K$ kernel is composed of $m$ use ports that are used to be connected to the $m$ quantities needed by the computation (\ie the numerical code). The component also exposes a provide port to be connected to the $Scheduler$ component. A generic $K$ component is represented in Fig.~\ref{fig:k}.

@@

Finally, the $Scheduler$ component is the component responsible for implementing the $TSP$ computed by MSC. Thus, this component represents the specific parallel orchestration of  computations. It exposes as many use ports as there are instances of $K$ components to call (\ie computations and fusions of computations). The component also exposes a provide port to be connected to the $Time$ component. A generic $Scheduler$ component is represented in Fig.~\ref{fig:ass}.

@@

\begin{figure}[t]
\begin{center}
\subfloat[][$K$\label{fig:k}]{
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$K$}; 
   \node[provide] (p) at (-1,0) {}; 
   \node[use] (u) at (1,0) {}; 
   \node[] (star) at (1,0.5) {$*$};
  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
}
\hspace{\fill}
\subfloat[][\label{fig:sync}$Sync$]{
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$Sync$};
   \node[provide] (p) at (-1,0) {};
   \node[use] (u) at (1,0) {$m$};

  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
}
\hspace{\fill}
\subfloat[][\label{fig:sched}$Scheduler$]{
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (sync) at (0,0) {$Scheduler$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[] (star) at (1.5,0.5) {$*$};
  \path[-]
    (p) edge node {} (sync)
    (sync) edge node {} (u);
\end{tikzpicture}
}
\caption{Specific components used to transform GA to the specialized component assembly of the simulation.}
\label{fig:specass}
\end{center}
\end{figure}

@@

   \node[provide,below right=0.2 of Du3] (Datap1) {};
   \node[component,draw=red,line width=0.4mm,fill=green!50!white,above=0.8cm of Datap2] (Data) {$A$};
   \node[component,draw=red,line width=0.4mm,fill=green!50!white,below=0.8cm of Data] (Data2) {$B$};
   \node[below=0.5cm of Data2] (Data3) {...};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[provide,right=1cm of Data] (mpidata) {};
   \node[right=0.2 of mpidata] (u1) {1};
   \node[provide,right=1cm of Data2] (mpidata3) {};
   \node[right=0.2 of mpidata3] (u2) {2};
   \node[provide,below=0.2cm of mpidata3] (mpidata2) {};
   \node[right=0.2 of mpidata2] (u3) {3};

   \node[provide,right=0.15 of Du1] (DDSp1) {};
   \node[provide,above=0.15 of Datau] (DDSp2) {};
   \node[component, draw=red, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {cart};

@@

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
        edge node {} (Du3)
    (DDSp1) edge node {} (DDS)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (scu1)
    	edge node {} (scu2)
    (Datap1) edge node {} (Data.west)
    		edge node {} (Data2.west)
    (Data) edge node {} (Datau)
          edge node {} (mpidata)
     (Data2) edge node {} (mpidata3)
     		edge node {} (mpidata2)
    (DDSp2) edge node {} (DDS)
    (pk0) edge node {} (k0)
    (psync1) edge node {} (sync1)
    (k0) edge node {} (k0u)
    	edge node {} (k0u2)
    (sync1) edge node {} (syncu)
    (Data2.west) edge [bend left] node {} (Datau);
    %(k0p.east) edge [bend right] node {} (mpidata.east)
    %(k0p2.east) edge [bend right] node {} (mpidata2.east)
    %(syncp.east) edge [bend right] node {} (mpidata3.east);
\end{tikzpicture}
\vspace*{.5em}
\caption{Generic Assembly according to the Multi-Stencil program formalism.
The complete assembly would be too complex to entirely draw and only a sub part is represented.
Some connections are represented by numbers instead of lines in order not to make the schema more clear.
\JB{etre plus clair sur les notations (Comme pour la Fig\ref{fig:ga}}}
\label{fig:specass}
\end{center}
\end{figure}

>>>>

\medskip
More formally, the data parallelization of a multi-stencil program of equation (\ref{eq:msp})
consists in a partitioning of the mesh $\mathcal{M}$ in $p$ sub-meshes $\mathcal{M}=\{\mathcal{M}_0,\dots,\mathcal{M}_{p-1}\}$.
This step can be performed by an external graph partitioner~\cite{Pellegrini:1996:SSP:645560.658570,DBLP:conf/ieeehpcs/HeleneS13,lachat:hal-00768916} and is handled by the DDS implementation of the third party HPC specialist.

@@

However, as a multi-stencil program is an iterative program, computations that happen after $k_j$ at the time iteration $t$ have also been computed before $k_j$ at the previous time iteration $t-1$. For this reason another case of synchronization has to be defined.

@@

%One can notice that both types of synchronizations are performed by all resources.
\begin{mydef}
The concatenation of two ordered lists of respectively $n$ and $m$ computations $l_1=[k_i]_{0 \leq i \leq n-1}$ and $l_2=[k'_i]_{0 \leq i \leq m-1}$ is denoted $l_1 \cdot l_2$ and is equal to a new ordered list $l_3=[k_0,\dots,k_{n-1},k'_0,\dots,k'_{m-1}]$.
\end{mydef}

@@

\paragraph{\textbf{Example}} Figure~\ref{fig:exmsl} gives an example of a $\mathcal{MSP}$ program. From this example, the following ordered list of computation kernels is extracted:
\begin{equation*}
\Gamma = [k_0,k_1,k_2,k_3,k_4,k_5,k_6,k_7,k_8]
\end{equation*}
From this ordered list of computation kernels $\Gamma$, and from the rest of the multi-stencil program, synchronizations can be automatically detected from the call to $F_{sync}(\Gamma,0)$ to get the synchronized ordered list of kernels:
\begin{equation}
\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_5,k_6,k_7,k_{7;8}^{sync},k_8],
\label{eq:exsync}
\end{equation}
\begin{subequations}
where
\begin{align}
        k_{0;1}^{sync}=(\emptyset,\{(B,nce)\},(B,\cup_{\phi \in D_1} nce(\phi))),\\
        k_{1;4}^{sync}=(\emptyset,\{(C,nec)\},(C,\cup_{\phi \in D_4} nec(\phi))),\\
        k_{7;8}^{sync}=(\emptyset,\{(I,ncc)\},(I,\cup_{\phi \in D_8} ncc(\phi))).
\end{align}
\end{subequations}

@@

These definitions are known as \emph{data hazards classification}. However, a specific condition on the computation domain, due the multi-stencils specific case, is introduced for the write after write case. One can note that the \texttt{independent} keyword of Fig.~\ref{fig:grammar} is useful in this case as the user explicitly indicates that $D_i \cap D_j=\emptyset$.

@@

One can note that $\Gamma_{dep}$ only takes into account a single time iteration. A complete dependency graph of the simulation could be built. This is a possible extension of this work.

@@

Valdes \& Al~\cite{Valdes:1979:RSP:800135.804393} have identified a forbidden shape, or sub-graph, called $N$, such that a DAG without this shape is MSPD.
%the following property is verified :
%\begin{myth}
%The transitive reduction of a DAG $G$ is MSPD if and only if it does not contain $N$ as a sub-graph.
%\end{myth}

Thus, as $\Gamma_{dep}$ is a DAG, by removing N-Shapes it is transformed to a MSPD. The intuition is illustrated in Fig.~\ref{fig:over}. Considering the figure without the dashed line, the sub graph forms a "N" shape. The fact is that this shape cannot be represented as a composition of sequences or parallel executions.
To remove such forbidden N-shapes of $\Gamma_{dep}=(V,E)$, we have chosen to apply an over-constraint with the relation $k_0 \prec k_3$, such that a complete bipartite graph is created for the sub-dag as illustrated in Figure~\ref{fig:over}. By adding this arc to the DAG, it is possible to identify its execution as $sequence(parallel(k_0;k_2);parallel(k_1;k3))$ represented by the TSP tree of Fig.~\ref{fig:solu}.

\begin{figure}[h!]
\begin{center}
\subfloat[Over-constraint on the forbidden $N$ shape.\label{fig:over}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (bc0) at (-1,0) {};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (0,-1) {$k_2$};
   \node[] (bc2) at (-1,-1) {};
   \node[] (c3) at (2,-1) {$k_3$};

  \path[->]
    (bc0) edge [dotted] node {} (c0)
    (bc2) edge [dotted] node {} (c2)
    (c0) edge node {} (c1)
          edge [dashed] node [swap] {} (c3)
    (c2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
}
\hspace{30pt}
\subfloat[TSP tree of Fig.~\ref{fig:over}\label{fig:solu}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (S0) at (0,0) {$\mathcal{S}$};
   \node[] (bS0) at (-1,-1) {};
   \node[] (P1) at (-1,1) {$\mathcal{P}$};
   \node[] (aP1) at (-1.5,2) {$k_0$};
   \node[] (aaP1) at (-0.5,2) {$k_2$};
   \node[] (P2) at (1,1) {$\mathcal{P}$};
   \node[] (c1) at (0.5,2) {$k_1$};
   \node[] (c3) at (1.5,2) {$k_3$};

  \path[->]
    (bS0) edge [dotted] node {} (S0)
    (S0) edge node {} (P1)
         edge node {} (P2)
    (P1) edge node [swap] {} (aP1)
          edge node [swap] {} (aaP1)
    (P2) edge node {} (c1)
        edge node {} (c3);
  \end{tikzpicture}
}
\label{fid:nshape}
\caption{N-Shape problem and solution}
\end{center}
\end{figure}

@@

Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization. Even though the MSL language is not dedicated to produce very optimized independent stencil codes, but to produce the parallel orchestration of computations, building the $TSP$ tree makes available an easy optimization when the data parallelization technique is the only one used. 
This optimization consists in proposing a valid merge of some computation kernels inside a single space loop. This is called a fusion. As previously explained in Section~\ref{sect:formalism}, MSL restrict the definition of a numerical computation by writing a single quantity at a time which avoids errors in manual fusion or counter-productive fusions for task parallelization. MSF guarantees that proposed fusions are correct and will not cause errors in the final results of the simulation.

@@

The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where the $parent(k)$ function returns the parent vertex of $k$ in the tree, and where $k_{i;j}^{fus}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$. Finally, $type(k)$ returns $comp$ if the kernel is a computation kernel, and $sync$  or $red$ otherwise.

@@

Third, an additional fusion case is possible and more tricky to find. Similarly to the first observation, two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are data dependent and what is written by the first one is read by the second one. The construction of the tree also guarantees that synchronizations are not needed between these computations, otherwise a $k^{sync}$ would have been inserted between them (inherited from $\Gamma_{sync}$). Thus, $w_i$ the quantity written by $k_i$ is common to these computations. Considering the following:
\begin{itemize}
\item $D_i \neq D_j$, which means that loop fusion is by default not possible,
\item $(r_j,n_j)$ is the pair read by $k_j$ for which $r_j=w_i$ and for which $n_j:D_j \rightarrow D_i^m$
\end{itemize}
the fusion of $k_i$ and $k_j$ is possible if and only if $\exists n:D_i \rightarrow D_j \in \mathcal{N}$ such that
\begin{equation*}
\bigcup_{\phi \in D_i} n(\phi) = D_j
\end{equation*}
This means that even if domains are different, a loop fusion is possible if an adequate neighborhood function can be found. One can note that this particular fusion case is equivalent to a \emph{scatter} optimization, often used when using unstructured meshes. One can also note that the computation $k_j$ will be written in a different manner if a scatter fusion is performed or not. This particular case is illustrated in Fig.~\ref{fig:fus3}.

@@

The developer will be notified of fusions in the output of MSC. This is not a problem by using MSF as the fusion is proposed before the developer actually write the numerical code of $k_j$.

@@

MSC takes a MSL file written using the grammar described in Section~\ref{sect:msl}, as well as the Generic Assembly presented in Fig.~\ref{fig:ga} as inputs, and generates a specialized component assembly that manages the parallel orchestration of the computations of the simulation. In this final assembly, that could be compared to a pattern or a skeleton of the simulation, the developer still has to fill-in the functions corresponding to the various computation kernels by using the DDS instantiation chosen into the specialized assembly.
The overall behavior of the compiler is as follows: %presented in Algorithm~\ref{alg:compiler}.
\begin{enumerate}
 \item it parses the MSL input file and generates $\Gamma$, the list of computation kernels,
 \item from $\Gamma$, it builds $\Gamma_{sync}$, the list including synchronizations for data parallelism using Algorithm $F_{sync}$ introduced in Section~\ref{sect:parallelism},
 \item from $\Gamma_{sync}$, it builds $\Gamma_{dep}$, the DAG supporting hybrid parallelism using Algorithm $F_{dep}$ introduced in Section~\ref{sect:parallelism},
 \item it then removes the N-Shapes from $\Gamma_{dep}$ to get a MSPD graph, and generates its serie-parallel binary tree decomposition $TSP$,
 \item it performs the fusion of kernels in $TSP$ if required (data parallelization only),
 \item it transforms GA to generate its output specialized component assembly.
\end{enumerate}

The last step of this compilation process is detailed below. It is composed of four steps:
\begin{enumerate}
 \item it instantiates $DDS$ and $Data$ components by using components implemented by a third party HPC specialist,
 \item it generates the structure of $K$ components responsible for each computation kernel of the simulation,
 \item it generates a new $Scheduler$ component,
 \item it replaces the $Computations$ component of GA by a generated sub-assembly that matches $TSP$ by using $Scheduler$, $K$ and $Sync$ components.
 \end{enumerate}

% detailed each component (interfaces)
New components have been introduced above and need to be explained. A $K$ component is a component into which the developer will write numerical code. It could represents a single computation kernel described by the numerician using MSL, or it could represents the fusion of multiple computation kernels. In any case the name of the generated component will use kernel identifiers used in the MSL description. A $K$ kernel is composed of $m$ use ports that are used to be connected to the $m$ quantities needed by the computation (\ie the numerical code). The component also exposes a provide port to be connected to the $Scheduler$ component. Interfaces of a $K$ component are represented in Fig.~\ref{fig:k}.

@@

Finally, the $Scheduler$ component is the component responsible for implementing the $TSP$ tree  computed by MSC. Thus, this component represents the specific parallel orchestration of  computations. It exposes as many use ports as there are instances of $K$ components to call (\ie computations and fusions of computations). The component also exposes a provide port to be connected to the $Time$ component. Interfaces of a $Scheduler$ component are represented in Fig.~\ref{fig:ass}.

@@

\begin{figure}[t]
\begin{center}
\subfloat[][$K$\label{fig:k}]{
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$K$}; 
   \node[provide] (p) at (-1,0) {}; 
   \node[use] (u) at (1,0) {}; 
   \node[] (star) at (1,0.2) {$*$};
  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
}
\hspace{\fill}
\subfloat[][\label{fig:sync}$Sync$]{
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (seq) at (0,0) {$Sync$};
   \node[provide] (p) at (-1,0) {};
   \node[use] (u) at (1,0) {$m$};

  \path[-]
    (p) edge node {} (seq)
    (seq) edge node {} (u);
\end{tikzpicture}
}
\hspace{\fill}
\subfloat[][\label{fig:sched}$Scheduler$]{
\begin{tikzpicture}[shorten >=1pt, node distance=1cm, on grid, auto]
   \node[component] (sync) at (0,0) {$Scheduler$};
   \node[provide] (p) at (-1.5,0) {};
   \node[use] (u) at (1.5,0) {};
   \node[] (star) at (1.5,0.2) {$*$};
  \path[-]
    (p) edge node {} (sync)
    (sync) edge node {} (u);
\end{tikzpicture}
}
\caption{Specific components used to transform GA to the specialized component assembly of the simulation.}
\label{fig:specass}
\end{center}
\end{figure}

@@

   \node[provide,below right=0.2 of Du3] (Datap1) {};
   \node[component,fill=red!50!white,above=0.8cm of Datap2] (Data) {$A$};
   \node[component, fill=red!50!white,below=0.8cm of Data] (Data2) {$B$};
   \node[below=0.5cm of Data2] (Data3) {...};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[provide,right=1cm of Data] (mpidata) {};
   \node[right=0.2 of mpidata] (u1) {1};
   \node[provide,right=1cm of Data2] (mpidata3) {};
   \node[right=0.2 of mpidata3] (u2) {2};
   \node[provide,below=0.2cm of mpidata3] (mpidata2) {};
   \node[right=0.2 of mpidata2] (u3) {3};

   \node[provide,right=0.15 of Du1] (DDSp1) {};
   \node[provide,above=0.15 of Datau] (DDSp2) {};
   \node[component, fill=red!50!white,above=0.8cm of DDSp2] (DDS) {cart};

@@

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
        edge node {} (Du3)
    (DDSp1) edge node {} (DDS)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (scu1)
    	edge node {} (scu2)
    (Datap1) edge node {} (Data.west)
    		edge node {} (Data2.west)
    (Data) edge node {} (Datau)
          edge node {} (mpidata)
     (Data2) edge node {} (mpidata3)
     		edge node {} (mpidata2)
    (DDSp2) edge node {} (DDS)
    (pk0) edge node {} (k0)
    (psync1) edge node {} (sync1)
    (k0) edge node {} (k0u)
    	edge node {} (k0u2)
    (sync1) edge node {} (syncu)
    (Data2.west) edge [bend left] node {} (Datau);
\end{tikzpicture}
\vspace*{.5em}
\caption{Sub-part of the specialized assembly generated by MSC from the example of the example of Fig.\ref{fig:exmsl} used throughout the paper. For readability some connections are represented by numbers instead of lines. The entire assembly is generated by MSC, however some components are automatically generated by MSC (in green), some are written by HPC specialists (in red) and others by the developer (in blue).}
\label{fig:specass}
\end{center}
\end{figure}
